---
title: "Main_analysis"
author: "Hillman Han"
date: '2022-09-12'
output: html_document
---
```{r}
#Load data and packages
library(readxl)
library(ggplot2)
library(dplyr)
library(corrplot)

df <- read.csv("train.csv")
df <- df[, 2: ncol(df)] #getting rid of id column
      
```

## Explore the dataset
Our dependent variable is SalePrice
```{r}
summary(df$SalePrice)

medianSP <- median(df$SalePrice)
ggplot(df, aes(x = SalePrice)) +
      geom_density() +
      scale_x_continuous(name = "Sales Price", breaks = seq(0,max(df$SalePrice), 1e+5)) +
      geom_vline(xintercept = medianSP, size = 1, color = "red") +
      geom_text(aes(x = medianSP + 50000, label = paste0("median\n", medianSP), y = 7e-06))
#This shows that a large proportion of sales price is centered around the median 163000

```
Finding numeric variables and categorical variables
```{r}
numV <- which(sapply(df, is.numeric)) #saving the indices of numeric columns
length(numV) #There are 37 variables that are numeric
numVnames <- names(numV)
```


## Explore whether the column has NA
```{r}
colSums(is.na(df)) #how many NAs in a column
NAcolumns <- names(which(colSums(is.na(df)) > 0 )) #which columns have NAs
print(NAcolumns)
cat('There are', length(names(which(colSums(is.na(df)) > 0 ))), 'columns with NAs. ')
cat('In these columns,', names(which(colSums(is.na(df)) > 0.8 * nrow(df))), ', more than 80% of total observations are NAs')

```
This shows that 19 columns contain missing values. 4 of these 19 columns  have more than 1168 missing values (80% of total obeservations). These 4 columns will be excluded from the list of independent variables of future models. The next step is to understand the data description and assign appropriate values to missing values for each column if needed, namely, LotFrontage, MasVnrType, MasVnrArea, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2, Electrical FireplaceQu, GarageType, GarageYrBlt, GarageFinish, GarageQual, GarageCond, PoolQC.

## LotFrontage
First explore the data using ggplot
```{r}

```






