---
title: "Main_analysis"
author: "Hillman Han"
date: '2022-09-12'
output: html_document
---
```{r}
#Load data and packages
library(readxl)
library(ggplot2)
library(dplyr)
library(corrplot)

df <- read.csv("train.csv")
df <- df[, 2: ncol(df)] #getting rid of id column
      
```

## Explore the dataset
Our dependent variable is SalePrice
```{r}
summary(df$SalePrice)

medianSP <- median(df$SalePrice)
ggplot(df, aes(x = SalePrice)) +
      geom_density() +
      scale_x_continuous(name = "Sales Price", breaks = seq(0,max(df$SalePrice), 1e+5)) +
      geom_vline(xintercept = medianSP, size = 1, color = "red") +
      geom_text(aes(x = medianSP + 50000, label = paste0("median\n", medianSP), y = 7e-06))
#This shows that a large proportion of sales price is centered around the median 163000

```
Finding numeric variables and categorical variables
```{r}
numV <- which(sapply(df, is.numeric)) #saving the indices of numeric columns
length(numV) #There are 37 variables that are numeric
numVnames <- names(numV)
charV <- which(sapply(df, is.character))
length(charV) #There are 43 variables that are categorical
charVnames <- names(charV)
charVnames
```


## Explore whether the column has NA
```{r}
colSums(is.na(df)) #how many NAs in a column
NAcolumns <- names(which(colSums(is.na(df)) > 0 )) #which columns have NAs
print(NAcolumns)
cat('There are', length(names(which(colSums(is.na(df)) > 0 ))), 'columns with NAs. ')
cat('In these columns,', names(which(colSums(is.na(df)) > 0.8 * nrow(df))), ', more than 80% of total observations are NAs')

```
This shows that 19 columns contain missing values. 4 of these 19 columns  have more than 1168 missing values (80% of total obeservations). These 4 columns will be excluded from the list of independent variables of future models. The next step is to understand the data description and assign appropriate values to missing values for each column if needed, namely, LotFrontage, MasVnrType, MasVnrArea, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2, Electrical FireplaceQu, GarageType, GarageYrBlt, GarageFinish, GarageQual, GarageCond, PoolQC.


## LotFrontage
Description: Linear feet of street connected to property.
Among all variables, it would make more sense if we use neighborhood as our predictor for the missing values in LotFrontage. The spread of LotFrontage for each neighborhood is relatively small. I am taking the median of LotFrontage for each neighborhood here.
```{r}
ggplot(df, aes(x = LotFrontage)) +
      geom_histogram() +
      facet_wrap(~Neighborhood)

df %>% 
      group_by(Neighborhood) %>%
      summarize(MedianLot = as.integer(median(LotFrontage, na.rm = TRUE))) %>%
ggplot(aes(x = as.factor(Neighborhood), y = MedianLot, fill = 'red')) +
      geom_col() +
      geom_text(aes(label = MedianLot), nudge_y = 0.5, vjust = -0.2, size = 3) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      ggtitle("median LotFrontage for each neighborhood") +
      xlab("Neighborhood") +
      theme(legend.position = 'none')

for (i in 1:nrow(df)){
        if(is.na(df$LotFrontage[i])){
               df$LotFrontage[i] <- as.integer(median(df$LotFrontage[df$Neighborhood == df$Neighborhood[i]], na.rm=TRUE)) 
        }
} #replace NAs in LotFrontage with their respective median from neiborhood group

#sanity check
sum(is.na(df$LotFrontage))
```

## MasVnrType & MasVnrArea
Description: Masonry veneer type; Masonry veneer area in square feet
Since for a house to have Masonry veneer area, it has to have a masonry veneer type. I need to check if the NA inputs in MasVnrType have inputs in MasVnr. If both have NAs, I will input none and 0 in the respective column.
```{r}
#See which rows have either NAs in MasVnrType or MasVnrArea
df[is.na(df$MasVnrType) | is.na(df$MasVnrArea), c('MasVnrType', 'MasVnrArea')] 

df$MasVnrType[is.na(df$MasVnrType)] <- 'none'
df$MasVnrArea[is.na(df$MasVnrArea)] <- 0

#Sanity Check
sum(is.na(df$MasVnrType))
sum(is.na(df$MasVnrArea))
```
## Bsmt_
Description:
BsmtQual: Evaluates the height of the basement
BsmtCond: Evaluates the general condition of the basement
BsmtExposure: Refers to walkout or garden level walls
BsmtFinType1: Rating of basement finished area
BsmtFinType2: Rating of basement finished area (if multiple types)
The logic to process basement related variables is the same as Masonry.
The description file tells that NA means no basement, but to make sure there is no missing input, I am checking there are rows where some of the five variables have inputs while others not. This will mean that NA is missing inputs instead of no basement.
```{r}
df[is.na(df$BsmtQual) | is.na(df$BsmtCond) | is.na(df$BsmtExposure) | is.na(df$BsmtFinType1) | is.na(df$BsmtFinType2), c('BsmtQual', 'BsmtCond', 'BsmtExposure', 'BsmtFinType1','BsmtFinType2')] 
#Two rows have NA for one of the five variables, while others have NAs for all five variables.
#Two rows are row 333 and row 949
ggplot(df, aes(x = BsmtFinSF2)) +
      geom_histogram() +
      facet_wrap(~BsmtFinType2) +
      ylim(0,5)
#There can be some correlation between BsmtFinSF2 and BsmtFinType2
df %>% 
      group_by(BsmtFinType2) %>%
      summarise(meanFinSF2 = mean(BsmtFinSF2))
df[333,] # has BsmtFinSF2 input 479, since I do not have a large sample for BsmtFinSF2, I can only take a bold step to match the input the closest mean, that is Rec
df[333,'BsmtFinType2'] <- 'Rec'

#For row 949, the NA is in BsmtExposure, I will replace it with the most common BsmtExposure
ggplot(df, aes(x = BsmtExposure)) +
      geom_bar()
df[949, 'BsmtExposure'] <- 'No'

#the other NAs will be replaced with 'NoBase'
df$BsmtQual[is.na(df$BsmtQual)] <- 'NoBase'
df$BsmtCond[is.na(df$BsmtCond)] <- 'NoBase'
df$BsmtExposure[is.na(df$BsmtExposure)] <- 'NoBase'
df$BsmtFinType1[is.na(df$BsmtFinType1)] <- 'NoBase'
df$BsmtFinType2[is.na(df$BsmtFinType2)] <- 'NoBase'

#sanity check
sum(is.na(df$BsmtQual))
sum(is.na(df$BsmtCond))
sum(is.na(df$BsmtExposure))
sum(is.na(df$BsmtFinType1))
sum(is.na(df$BsmtFinType2))

```


## Garage_
Description:
GarageType: Garage location
GarageYrBlt: Year garage was built
GarageFinish: Interior finish of the garage
GarageQual: Garage quality
GarageCond: Garage condition
The logic to process basement related variables is the same as Masonry and Basement. Since the other two garage related variables, namely GarageCars and GarageArea, do not have missing inputs, I will cross check with these two columns first. 
```{r}
df[is.na(df$GarageType) | is.na(df$GarageYrBlt) | is.na(df$GarageFinish) | is.na(df$GarageQual) | is.na(df$GarageCond), c('GarageType', 'GarageYrBlt', 'GarageFinish', 'GarageQual','GarageCond', 'GarageCars', 'GarageArea')] 
#It turns out all NA here match GarageCars = 0 and GarageArea = 0, meaning No Garage in the house. Therefore, I move on to replace all NAs with 'NoGar' to clarify the category.

df$GarageType[is.na(df$GarageType)] <- 'NoGar'
df$GarageYrBlt[is.na(df$GarageYrBlt)] <- 'NoGar'
df$GarageFinish[is.na(df$GarageFinish)] <- 'NoGar'
df$GarageQual[is.na(df$GarageQual)] <- 'NoGar'
df$GarageCond[is.na(df$GarageCond)] <- 'NoGar'

#Sanity Check
sum(is.na(df$GarageType))
sum(is.na(df$GarageYrBlt))
sum(is.na(df$GarageFinish))
sum(is.na(df$GarageQual))
sum(is.na(df$GarageCond))

```

## Electrical
Description: Electrical system
There is only 1 NA in this column, I will use the most frequent of all types.
```{r}
ggplot(df, aes(x = Electrical)) +
      geom_bar()

df$Electrical[is.na(df$Electrical)] <- 'SBrkr'
sum(is.na(df$Electrical))
```

## FireplaceQu
Description: Fireplace quality
My assumption is the NAs in this column mean that there is no Fireplace in the house. It would be then be reasonable that Fireplaces = 0; otherwise, if the Fireplaces >0, the NAs in DireplaceQu are actually missing inputs. I will use the most common type to replace the NA.
```{r}
sum((is.na(df$FireplaceQu)) & (df$Fireplaces == 0)) == sum(is.na(df$FireplaceQu))
#All NAs in FireplaceQu mean there is no fireplace in the house.

df$FireplaceQu[is.na(df$FireplaceQu)] <- 'NoFire'
sum(is.na(df$FireplaceQu))
```

## Reassemble dataframe
```{r}
train <- df %>%
      select(-Alley, -PoolQC, -Fence, -MiscFeature)
sum(is.na(train)) #clean
```













